\documentclass{moderncv}

% Style
\moderncvtheme[blue]{classic}
\usepackage{multicol}

% Document Language
\usepackage[english]{babel}

% Font
\ifxetexorluatex
  \usepackage{fontspec}
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX}
  \setmainfont{Latin Modern Roman}
  \setsansfont{Latin Modern Sans}
  \setmonofont{Latin Modern Mono}
  \setmathfont{Latin Modern Math} 
\else
  \usepackage[T1]{fontenc}
  \usepackage{lmodern}
\fi

% Scale
\usepackage[scale=0.85, top=1.5cm, bottom=2cm]{geometry}
\setlength{\hintscolumnwidth}{2.7cm}

% Personal Informations
\name{}{Samy Lahlou Kamal}
\phone[mobile]{+1~(514)~699~4404}
\email{samy.lahloukamal@mcgill.ca}  
\homepage{samylahlou.com} 

\renewcommand*{\bibliographyitemlabel}{[\arabic{enumiv}]}

\begin{document}

\makecvtitle

\vspace{-5mm}

% EDUCATION
\section{Education}
\cvitemwithcomment{August 2022 \qquad -- present}{\textbf{McGill University}}{\normalfont{\textbf{Montréal, Canada}}}
\cvitem{}{\vspace{-9.4mm}
Bachelor of Science - Honours Mathematics \& Minor Computer Science.}
\vspace{-1mm}
\cvitem{}{\vspace{-9.4mm} GPA: 3.79}
\vspace{-4mm}

\cvitemwithcomment{September 2015 -- June 2022}{\textbf{Lycée Lyautey}}{\normalfont{\textbf{Casablanca, Morocco}}}
\cvitem{}{\vspace{-9.4mm} French Baccalaureate with High Honors (Mention Très Bien).}
\vspace{-1mm}
\cvitem{}{\vspace{-9.4mm} Specialities: Mathematics \& Computer Science}
\vspace{-1mm}
\cvitem{}{\vspace{-9.4mm} Options: Expert Mathematics \& Music}
\vspace{-6.5mm}

%RESEARCH EXPERIENCE
\section{Research Experience}
\cvitemwithcomment{May 2025 \qquad \quad-- present}{\textbf{Research Project on Special Values of L-functions}}{}
\cvitem{}{\vspace{-9.35mm} Supervisor: Professor \href{https://en.wikipedia.org/wiki/Henri_Darmon}{Henri Darmon} (McGill University)}
\cvitem{}{ \vspace{-9.8mm}
\begin{itemize}
    \item Currently working on the history of L-functions (reading the original books and papers), and their special values.
    \item Presented my work in the \href{https://www.math.mcgill.ca/darmon/courses/24-25/usra/usra.html}{USRA Number Theory seminar}.
    \item Started writing an expository paper on the detailed history of L-functions and the mathematicians that contributed to that subject (L. Euler, P.L. Dirichlet, B. Riemann, R. Dedekind, E. Hecke, ...): [\href{https://samylahlou.com/NSERC/Report/nserc_report_samylahlou.pdf}{pdf (unfinished)}]
    \item Funded by the \href{https://www.nserc-crsng.gc.ca/}{NSERC (Natural Sciences and Engineering Research Council of Canada)}.
\end{itemize}
}
\vspace{1.5mm}

\cvitemwithcomment{Summer 2024}{\textbf{Research Project on the Ergodic Theorems}}{}
\cvitem{}{\vspace{-4.1mm} Supervisor: Professor \href{https://www.math.mcgill.ca/atserunyan/}{Anush Tserunyan} (McGill University)}
\cvitem{}{ \vspace{-4.5mm}
\begin{itemize}
    \item Studied Measured Group Theory with applications to the Ergodic Theorems.
    \item Wrote an expository paper on the ergodic theorems with a focus on Pr. Tserunyan's new proof of the Birkhoff Ergodic Theorem: [\href{https://samylahlou.com/papers/MATH470/samy_ergodic_report.pdf}{pdf}]
    \item This project was part of the Honours Research Project class MATH 470.
\end{itemize}
}
\vspace{1.5mm}

\cvitemwithcomment{January 2024 \quad -- April 2024}{\textbf{Directed Reading Program on $p \, $}-\textbf{adic Numbers}}{}
\cvitem{}{\vspace{-9.35mm} Supervisor: Doctoral student \href{https://www.math.mcgill.ca/hazem.hassan/}{Hazem Hassan} (McGill University)}
\cvitem{}{ \vspace{-9.8mm}
\begin{itemize}
    \item Studied $p$-adic numbers and $p$-adic Zeta-Functions by reading the book \textit{$p$-adic Numbers, $p$-adic Analysis, and Zeta-Functions} by Neal Koblitz.
    \item Presented my learnings in front the other \href{https://www.math.mcgill.ca/gsams/drp/index.html}{Directed Reading Program} students.
    \item Wrote a report presenting $p$-adic numbers, their construction and $p$-adic interpolation.
\end{itemize}
}
\vspace{-0mm}

% PROFESSIONAL EXPERIENCE
\section{Professional Experience}

\cvitemwithcomment{September 2025 -- present}{\textbf{Teaching Assistant for the course History \& Philosophy of Mathematics}}{}
\vspace{-6mm}
\cvitem{}{\textit{McGill University, Montréal, QC}}
\vspace{-0.8mm}
\cvitem{}{Instructor: Professor Henri Darmon, Course Code: MATH 338, 135 students}
\vspace{-0.8mm}
\cvitem{}{
\begin{itemize}
    \item Conducted tutorial sessions (three hours a week) and exam review sessions to go over the course content, additional content, and answer students' questions. 
    \item Wrote a suitable set of notes based on the students' questions to complete the ressources available to the students.
    \item Planned one-on-one meetings with students who wanted to.
\end{itemize}
}
\vspace{1.5mm}

\cvitemwithcomment{January 2024 \quad -- April 2024}{\textbf{Grader for the course Discrete Structures}}{}
\vspace{-6mm}
\cvitem{}{\textit{McGill University, Montréal, QC}}
\vspace{-0.8mm}
\cvitem{}{Instructor: Mr. Jeremy Macdonald, Course Code: MATH 240, 365 students}
\vspace{-0.8mm}
\cvitem{}{
\begin{itemize}
    \item Graded assignments (one every two weeks) for the course Discrete Structures.
\end{itemize}
}
\vspace{1.5mm}

\cvitemwithcomment{May 2023 -- September 2024}{\textbf{Tutor in Mathematics}}{}
\vspace{-6mm}
\cvitem{}{\textit{McGill University, Montréal, QC}}
\vspace{-0.8mm}
\cvitem{}{
\begin{itemize}
    \item Planning tutoring sessions (private or in group) for 1st year students in Mathematics.
\end{itemize}
}
\vspace{1.5mm}

\cvitemwithcomment{June 2024 \qquad \quad -- present}{\textbf{Busboy}}{}
\vspace{-6mm}
\cvitem{}{\textit{Cirque Eloize, Montréal, QC}}
\vspace{-0.8mm}
\cvitem{}{
\begin{itemize}
    \item Supported bartenders during large events by restocking beverages, supplies, and cleaning.
\end{itemize}
}
\vspace{1.5mm}

\cvitemwithcomment{May 2024 \qquad \quad -- August 2024}{\textbf{Busboy}}{}
\vspace{-6mm}
\cvitem{}{\textit{Pizzeria Moretti Griffintown, Montréal, QC}}
\vspace{-0.8mm}
\cvitem{}{
\begin{itemize}
    \item Supported servers by cleaning tables, resetting place seetings, refilling water and supplies, and maintaining clean dining environment.
\end{itemize}
}

%PROJECTS
\section{Projects}

\cvitemwithcomment{January 2025}{\textbf{Expository Paper on the Divergence of the Harmonic Series}}{}
\vspace{-0.5mm}
\cvitem{}{
\begin{itemize}
    \item Wrote a self-contained paper presenting the importance of rigor in mathematics through the detailed study of the divergence of the Harmonic Series: [\href{https://samylahlou.com/papers/HarmonicSeries/FullVersion/harmonic_series_samy.pdf}{pdf}]
\end{itemize}
}
\vspace{1.5mm}

\cvitemwithcomment{December 2024}{\textbf{Published Historical Article on Fourier Analysis}}{}
\vspace{-0.5mm}
\cvitem{}{
\begin{itemize}
    \item Researched on the origins and impacts of Fourier analysis on modern mathematics through the original works of the great mathematicians: J. Fourier, P.L. Dirichlet, B. Riemann, G. Cantor.
    \item Co-authored (along with Nisrine Sqalli) a historical paper presenting our findings: [\href{https://samylahlou.com/papers/Fourier/Fourier_Paper_Full.pdf}{pdf}]
    \item Published a condensed version in the \href{https://thedeltaepsilonmcgill.github.io/}{Delta Epsilon journal}: [\href{https://thedeltaepsilonmcgill.github.io/pdfs/DE8.pdf}{pdf}]
\end{itemize}
}
\vspace{1.5mm}

\cvitemwithcomment{December 2024 \quad -- present}{\textbf{Personal Website}}{}
\cvitem{}{ \vspace{-9mm}
\begin{itemize}
    \item Coded a personal website by learning and using HTML and CSS: [\href{https://samylahlou.com}{samylahlou.com}]
    \item Post regularly my personal solutions to textbook exercises, articles that I write and reviews of textbooks and books related to mathematics.
\end{itemize}
}

% AWARDS
\section{Awards}
\cvitem{Summer 2025}{\textbf{NSERC Undergraduate Student Research Award -- \$8950}}
\vspace{1mm}

% Coding Skills
\section{Coding Skills}
\begin{cvcolumns}
    \cvcolumn{Programing Languages}{
        \vspace{-8mm}
        \begin{multicols}{2}
        \begin{itemize}
            \item Python
            \item Java
            \item C
            \item Bash
            \item Assembly
            \item SQL
            \item MATLAB
            \item SageMath
        \end{itemize}
        \end{multicols}}
    \cvcolumn{Markup Languages}{
        \begin{itemize}
            \item \LaTeX
            \item HTML
            \item CSS
        \end{itemize}}
\end{cvcolumns}

% LANGUAGES
\section{Languages}
\begin{cvcolumns}
    \cvcolumn{}{English (Fluent)}
    \cvcolumn{}{French (Native)}
    \cvcolumn{}{Arabic (Native)}
\end{cvcolumns}


\end{document}